---
title: "R Notebook"
output: html_notebook
---


Trend analysis of leatherback turtles nesting in Florida, Puerto Rico, and St Croix with Kelly Stewart.

New data were received in August 2020. 

```{r}
rm(list=ls())
library(tidyverse)
library(ggplot2)
library(readr)
library(lubridate)
library(jagsUI)
library(bayesplot)
library(stringr)

MCMC.params <- list(n.samples = 100000,
                    n.burnin = 50000,
                    n.thin = 5,
                    n.chains = 5)


```

Get data;

```{r}
col.def <- cols(ID = col_integer(),
                year = col_integer(),
                beach = col_character(),
                latitude = col_double(),
                distance = col_double(),
                days_week = col_integer(),
                days_year = col_integer(),
                nests = col_integer())

FL.nest.counts <- read_csv("data/FL_Sept2020.csv", 
                           col_types = col.def) %>% 
  mutate(beach_f = as.factor(toupper(beach)),
         dataset = "FL",
         ID2 = as.numeric(as.factor(ID)))

PR.nest.counts <- read_csv("data/PR_Sept2020.csv", 
                           col_types = col.def) %>% 
  mutate(beach_f = as.factor(toupper(beach)),
         dataset = "PR",
         ID2 = as.numeric(as.factor(ID)))

STX.nest.counts <- read_csv("data/STX_Sept2020.csv", 
                           col_types = col.def) %>% 
  mutate(beach_f = as.factor(toupper(beach)),
         dataset = "STX",
         ID2 = as.numeric(as.factor(ID)))


```


Take a look at each area. There are so many beaches in FL and difficult to see them all. 

```{r}
p.FL <- ggplot(data = FL.nest.counts) +
  geom_path(aes(x = year, y = log(nests),
            color = factor(ID))) +
  theme(legend.position = "none")
  #facet_wrap( ~ factor(ID))

p.FL

```



```{r}
p.PR <- ggplot(data = PR.nest.counts) +
  geom_path(aes(x = year, 
                y = log(nests),
                color = factor(ID)))+
  theme(legend.position = "none")

p.PR
```



```{r}
p.STX <- ggplot(data = STX.nest.counts) +
  geom_path(aes(x = year, y = log(nests)))

p.STX
```

Try one dataset at a time. 

For FL, we can fit the same model as the last publication with 3 covariates. 

```{r}
X.FL <- cbind(FL.nest.counts$distance - median(FL.nest.counts$distance),
                FL.nest.counts$days_week - median(FL.nest.counts$days_week),
                FL.nest.counts$days_year - median(FL.nest.counts$days_year))

FL.nest.counts %>% 
  group_by(ID2) %>%
  summarise(n = n()) -> FL.ns

FL.nest.counts %>% 
  select(ID2, ID, latitude, beach_f, year) %>% 
  group_by(ID) %>%
  summarise(ID2 = first(ID2),
            name = first(beach_f),
            latitude = first(latitude)) %>%
  mutate(latc = latitude - mean(latitude),
         latc2 = latc^2,
         beach = ID,
         ID2 = ID2,
         name = name,
         n = FL.ns$n) -> FL.lat.dat

FL.jags.data <- list(N = length(FL.nest.counts$nests),
                     nbeach = length(unique(FL.nest.counts$ID2)),
                     count = FL.nest.counts$nests,
                     beach = FL.nest.counts$ID2,
                     yearc = FL.nest.counts$year - median(FL.nest.counts$year),
                     X = X.FL,
                     latc = FL.lat.dat$latc,
                     latc2 = FL.lat.dat$latc2)
```

Run the model for FL

```{r}

parameters <- c("a0", "a1", "beta", "B.hat", "a1pc",
                "mu.a0", "mu.a1",
                "rea0", "rea1",
                "delta1", "delta2",
                "sigma.e", "Sigma.B",
                "sigma.a0", "sigma.a1",
                "floridapc", "probP", "deviance",
                "Devobs", "Devpred")

MCMC.params$model.file = "models/Model_JAGS_rSlope_rInt_3Covs.txt"

if (!file.exists("RData/JAGS_out_rSlope_rInt_3Covs_FL.rds")){
  jm.FL <- jags(data = FL.jags.data,
                #inits = inits,
                parameters.to.save= parameters,
                model.file = MCMC.params$model.file,
                n.chains = MCMC.params$n.chains,
                n.burnin = MCMC.params$n.burnin,
                n.thin = MCMC.params$n.thin,
                n.iter = MCMC.params$n.samples,
                DIC = T, 
                parallel=T)
  
  saveRDS(jm.FL, 
          file = "RData/JAGS_out_rSlope_rInt_3Covs_FL.rds")
  
} else {
  jm.FL <- readRDS("RData/JAGS_out_rSlope_rInt_3Covs_FL.rds")
}

summary(jm.FL)
```





STX

```{r}
# The X matrix for STX is not useful because there is only one beach, except the number of days per year surveyed. 
X.STX <- cbind(STX.nest.counts$distance - median(STX.nest.counts$distance),
                STX.nest.counts$days_week - median(STX.nest.counts$days_week),
                STX.nest.counts$days_year - median(STX.nest.counts$days_year))

STX.jags.data <- list(N = length(STX.nest.counts$nests),
                      nbeach = length(unique(STX.nest.counts$ID2)),
                      count = STX.nest.counts$nests,
                      beach = STX.nest.counts$ID2,
                      yearc = (STX.nest.counts$year - median(STX.nest.counts$year)),
                      X = X.STX)

                      # latc = STX.lat.dat$latc,
                      # latc2 = STX.lat.dat$latc2)


```

The original model doesn't work for just one beach. I can add STX and PR to the FL dataset and analyze them all at once. Or 
Then run the model for STX. Note that some covariates in the original model are unavailable. I changed it. 

```{r}
parameters <- c("a0", "a1", "beta", "B.hat", "a1pc",
                "mu.a0", "mu.a1",
                "rea0", "rea1",
                "delta1", "delta2",
                "sigma.e", "Sigma.B",
                "sigma.a0", "sigma.a1",
                "floridapc", "probP", "deviance",
                "Devobs", "Devpred")

MCMC.params$model.file = "models/Model_JAGS_rSlope_rInt_3Covs.txt"



if (!file.exists("RData/JAGS_out_Dist_rSlope_rInt_3Covs_2008.rds")){
  jm.2008 <- jags(data = jags.data.2008,
             #inits = inits,
             parameters.to.save= parameters,
             model.file = MCMC.params$model.file,
             n.chains = MCMC.params$n.chains,
             n.burnin = MCMC.params$n.burnin,
             n.thin = MCMC.params$n.thin,
             n.iter = MCMC.params$n.samples,
             DIC = T, 
             parallel=T)
  
  saveRDS(jm.2008, file = "RData/JAGS_out_Dist_rSlope_rInt_3Covs_2008.rds")
  
} else {
  jm.2008 <- readRDS("RData/JAGS_out_Dist_rSlope_rInt_3Covs_2008.rds")
}

summary(jm.2008)
```

At least they converged fine! 

Note that beach IDs are ID2, not the original ID numbers. So, we need to make sure right ones are used. 

```{r}
summary.2008.df <- data.frame(jm.2008$summary) %>%
  rownames_to_column("Parameter")

# find the beach-specific growth rate in percent
#mcmc_trace(jm.MS$samples, pars = "a1pc[38]")

summary.2008.df %>% filter(str_detect(Parameter, pattern = "a1pc")) %>%
  rownames_to_column("beach") %>%
  transmute(ID = as.numeric(beach),
            mean_a1pc = mean,
            sd_a1pc = sd,
            low25_a1pc = X2.5.,
            median_a1pc = X50.,
            upper975_a1pc = X97.5.)-> a1pc.2008

a1pc.2008 %>% left_join(lat.dat.2008, by = "ID") %>%
  arrange(latc) -> a1pc.2008

summary.2008.df %>% 
  filter(str_detect(Parameter, pattern = "mu.a1")) %>%
  rownames_to_column("beach") %>%
  transmute(ID = as.numeric(beach),
            mean_mu.a1 = mean,
            sd_mu.a1 = sd,
            low25_mu.a1 = X2.5.,
            median_mu.a1 = X50.,
            upper975_mu.a1 = X97.5.)-> mu.a1.2008

mu.a1.2008 %>% left_join(lat.dat.2008, by = "ID") %>%
  arrange(latc) -> mu.a1.2008

#head(a1pc.2008)
```


Look at why a1pc.2008 is quite different from a1pc.new...

Compare results visually...
```{r}

ggplot() + 
  geom_point(data = a1pc, 
             aes(x = beach, y = mean_a1pc),
             color = "blue") + 
  geom_errorbar(data = a1pc,
                aes(ymin = low25_a1pc,
                    ymax = upper975_a1pc,
                    x = beach),
                color = "blue") + 
  geom_point(data = a1pc.new, 
             aes(x = beach, y = mean_a1pc),
             color = "red") + 
  geom_errorbar(data = a1pc.new,
                aes(ymin = low25_a1pc,
                    ymax = upper975_a1pc,
                    x = beach),
                color = "red") + 
  geom_point(data = a1pc.2008, 
            aes(x = ID, y = mean_a1pc),
            color = "green") + 
  geom_errorbar(data = a1pc.2008,
              aes(ymin = low25_a1pc,
                  ymax = upper975_a1pc,
                  x = ID),
              color = "green")

```
