---
title: "R Notebook"
output: html_notebook
---

Trend analysis of leatherback turtles nesting in Florida, Puerto Rico, and St Croix with Kelly Stewart. IN this version (v.2), I use a state-space model that we used for loggerheads/leatherbacks for HI shallow-set longline biological review. The model is similar to ones in Boyd et al. (yr?)

```{r}
rm(list=ls())
library(tidyverse)
library(ggplot2)
library(readr)
library(lubridate)
library(jagsUI)
library(bayesplot)
```

Get data;

```{r}
col.def <- cols(ID = col_integer(),
                year = col_integer(),
                beach = col_character(),
                latitude = col_double(),
                distance = col_double(),
                days_week = col_integer(),
                days_year = col_integer(),
                nests = col_integer())

nest.counts <- read_csv("data/FL_PR_STX.csv", col_types = col.def) %>% 
  mutate(lat_band = ifelse(latitude < 26, 25,
                           ifelse(latitude < 27, 26,
                                  ifelse(latitude < 28, 27,
                                         ifelse(latitude < 29, 28,
                                                ifelse(latitude < 30, 29,
                                                       ifelse(latitude < 31, 30)))))),
         beach_f = as.factor(toupper(beach)))

```

Time series are different lengths. It'd be good to get posteriors on missing observations too. Rather than getting all ys, specify which ones are needed. 

```{r}
# There are some skipped years, which needs to be filled in
nest.counts %>% group_by(ID) %>%
  summarise(n.years = max(year) - min(year) + 1,
            year.1 = min(year),
            year.2 = max(year)) -> summary.years

year.mat <- y <- matrix(nrow = nrow(summary.years), 
                        ncol = max(summary.years$n.years))

# find which data points are missing.
NA.idx <- vector(mode = "list", length = nrow(summary.years))

k <- 3
for (k in 1:nrow(summary.years)){
  nest.counts %>% filter(ID == summary.years$ID[k]) %>%
    select(year, nests) %>% 
    mutate(seq.yr = year - min(year) + 1) -> tmp
  y[k,tmp$seq.yr] <- log(tmp$nests + 1)
  year.mat[k, tmp$seq.yr] <- tmp$year
  n.years.k <- summary.years$n.years[k]
  if (sum(is.na(year.mat[k, 1:n.years.k])) > 0)
    NA.idx[[k]] <- c(1:n.years.k)[is.na(year.mat[k, 1:n.years.k])]
}
```

Try fitting growth models - In this version, we use the state-space approach that is the same as one from HI SSLL BiOp. We look at the predicted missing values also. 

```{r}
jags.data <- list(n.beaches = nrow(y),
                  n.years = summary.years$n.years,
                  y = y)

MCMC.params <- list(n.samples = 100000,
                    n.burnin = 50000,
                    n.thin = 5,
                    n.chains = 5)

missing.y <- c()

for (k in 1:length(NA.idx)){
  if (length(NA.idx[[k]]) > 0){
    for (k1 in 1:length(NA.idx[[k]])){
      missing.y <- c(missing.y, paste0("y[", k, ",", NA.idx[[k]][k1], "]"))
    }
  }
}

parameters <- c(c("U", "mean.U", "sigma.U", 
                "sigma.Q", "sigma.R", "loglik"),
                missing.y)

MCMC.params$model.file = "models/Model_norm_norm_UsQsRs.txt"

if (!file.exists("RData/JAGS_out_norm_norm_UsQsRs.rds")){
  jm <- jags(data = jags.data,
             #inits = inits,
             parameters.to.save= parameters,
             model.file = MCMC.params$model.file,
             n.chains = MCMC.params$n.chains,
             n.burnin = MCMC.params$n.burnin,
             n.thin = MCMC.params$n.thin,
             n.iter = MCMC.params$n.samples,
             DIC = T, 
             parallel=T)
  
  saveRDS(jm, file = "RData/JAGS_out_norm_norm_UsQsRs.rds")
  
} else {
  jm <- readRDS("RData/JAGS_out_norm_norm_UsQsRs.rds")
}

summary(jm)

#mcmc_trace(jm$samples, pars = c("U[1]", "U[2]", "U[3]"))
#mcmc_dens(jm$samples, pars = c("y[8,3]", "y[8,4]", "y[8,5]"))
# mcmc_trace(jm$samples, pars = c("delta1", "delta2"))
# mcmc_dens(jm$samples, pars = c("delta1", "delta2"))

```

These parameters look good and all Rhat stats are < 1.1.  

```{r}
summary.df <- data.frame(jm$summary) %>% rownames_to_column(var = "parameter") 

```

For each beach, overlay data and estimated population growth +/- 95% CI.

Plot each beach 
```{r}
p <- 28
for (p in 1:length(NA.idx)){
  ID.1 <- p
  nest.counts %>% filter(ID == ID.1) -> nest.counts.ID
  
  summary.df %>% filter(parameter == paste0("U[", ID.1, "]")) %>%
    select(X2.5., X50., X97.5.)-> post.stats.U.ID 
  
  log.N0 <- log(nest.counts.ID$nests[1] + 1)
  log.Nt <- log.Nt.low <- log.Nt.high <- vector(mode = "numeric", 
                                                length = nrow(nest.counts.ID))
  log.Nt[1] <- log.Nt.low[1] <- log.Nt.high[1] <- log.N0
  for (k in 2:nrow(nest.counts.ID)){
    log.Nt[k] <- log.Nt[k-1] + post.stats.U.ID$X50. 
    log.Nt.low[k] <- log.Nt.low[k-1] + post.stats.U.ID$X2.5. 
    log.Nt.high[k] <- log.Nt.high[k-1] + post.stats.U.ID$X97.5. 
  }
  
  # convert them into anti-log space
  Nt <- exp(log.Nt) - 1
  Nt.low <- exp(log.Nt.low) - 1
  Nt.high <- exp(log.Nt.high) - 1
  nest.counts.ID$Nt <- Nt
  nest.counts.ID$Nt.low <- Nt.low
  nest.counts.ID$Nt.high <- Nt.high
  
  p.1 <- ggplot() + 
    geom_point(data = nest.counts.ID,
               aes(x = year, y = nests),
               color = "black") + 
    geom_point(data = nest.counts.ID,
               aes(x = year, y = Nt),
               color = "gold") + 
    geom_ribbon(data = nest.counts.ID,
                aes(x = year, ymin = Nt.low, ymax = Nt.high),
                fill = "gold", alpha = 0.5) + 
    labs(title = paste0("ID = ", ID.1)) + 
    xlab(element_blank()) + 
    ylab("Nests")
  
  # find out missing data
  if (length(NA.idx[[ID.1]]) > 0){
    pred.y <- summary.df[grep(paste0("y[", ID.1, ","), 
                              summary.df$parameter, fixed = T),
                         c("X2.5.", "X50.", "X97.5.")]
    missing.year <- min(year.mat[ID.1]) + NA.idx[[ID.1]] - 1
    pred.y.df <- data.frame(year = missing.year,
                            y.med = max((exp(pred.y$X50.) - 1), 0),
                            y.low = max((exp(pred.y$X2.5.) - 1), 0),
                            y.high = (exp(pred.y$X97.5.) - 1))
    p.1 <- p.1 + 
      geom_point(data = pred.y.df,
                 aes(x = year, y = y.med),
                 color = "green") + 
      geom_errorbar(data = pred.y.df,
                    aes(x = year, ymin = y.low, ymax = y.high),
                    color = "green")
  }
  
  ggsave(p.1, filename = paste0("figures/beach_", p, ".png"),
         device = "png", dpi = 600)
}

```

