---
title: "R Notebook"
output: html_notebook
---

Trend analysis of leatherback turtles nesting in Florida, Puerto Rico, and St Croix with Kelly Stewart.

```{r}
rm(list=ls())
library(tidyverse)
library(ggplot2)
library(readr)
library(lubridate)
library(jagsUI)
library(bayesplot)
library(stringr)

MCMC.params <- list(n.samples = 100000,
                    n.burnin = 50000,
                    n.thin = 5,
                    n.chains = 5)

parameters <- c("a0", "a1", "beta", "B.hat", "a1pc",
                "mu.a0", "mu.a1",
                "rea0", "rea1",
                "delta1", "delta2",
                "sigma.e", "Sigma.B",
                "sigma.a0", "sigma.a1",
                "floridapc", "probP", "deviance",
                "Devobs", "Devpred")

```

Try fitting growth models - we can use the same model as the last paper as required/requested by KS. 

Get data;

```{r}
col.def <- cols(ID = col_integer(),
                year = col_integer(),
                beach = col_character(),
                latitude = col_double(),
                distance = col_double(),
                days_week = col_integer(),
                days_year = col_integer(),
                nests = col_integer())

nest.counts.FL <- read_csv("data/FL_07262020_FINAL.csv", 
                           col_types = col.def) %>% 
  mutate(beach_f = as.factor(toupper(beach)))

max.ID <- max(nest.counts.FL$ID)

nest.counts.PR <- read_csv("data/PR_07262020_FINAL.csv", 
                           col_types = col.def) %>% 
  mutate(beach_f = as.factor(toupper(beach)),
         ID = ID + max.ID,
         latitude = 18.467761)

max.ID <- max(nest.counts.PR)
nest.counts.STX <- read_csv("data/STX_07262020_FINAL.csv", 
                            col_types = col.def) %>% 
  mutate(beach_f = as.factor(toupper(beach)))

nest.counts.STX$latitude <- 17.679899

nest.counts <- rbind(nest.counts.FL, nest.counts.PR, nest.counts.STX) %>%
  mutate(ID2 = as.numeric(as.factor(ID))) -> nest.counts

nest.counts %>% 
  group_by(ID2) %>%
  summarise(n = n()) -> nest.counts.ns

nest.counts %>% 
  select(ID2, ID, latitude, beach_f) %>% 
  group_by(ID) %>%
  summarise(latitude = first(latitude),
            ID2 = first(ID2),
            name = first(beach_f)) %>%
  mutate(latc = latitude - mean(latitude),
         latc2 = latc^2,
         beach = ID,
         ID2 = ID2,
         name = name,
         n = nest.counts.ns$n) -> lat.dat 

# length surveyed, # days per week surveyed, # days surveyed per year
# 
X <- cbind(nest.counts$distance - median(nest.counts$distance), 
           nest.counts$days_week - median(nest.counts$days_week),
           nest.counts$days_year - median(nest.counts$days_year))

median.year <- round(median(seq(min(nest.counts$year),
                                max(nest.counts$year))))

# jags.data.2008 contains wrong counts... check this!
jags.data <- list(N = length(nest.counts$nests),
                  nbeach = length(unique(nest.counts$ID2)),
                  count = nest.counts$nests,
                  beach = nest.counts$ID2,
                  yearc = (nest.counts$year - median.year),
                  X = X,
                  latc = lat.dat$latc,
                  latc2 = lat.dat$latc2)
```

Then run the model:

```{r}
MCMC.params$model.file = "models/Model_JAGS_rSlope_rInt_3Covs.txt"

if (!file.exists("RData/JAGS_out_Dist_rSlope_rInt_3Covs_Aug2020.rds")){
  jm <- jags(data = jags.data,
             #inits = inits,
             parameters.to.save= parameters,
             model.file = MCMC.params$model.file,
             n.chains = MCMC.params$n.chains,
             n.burnin = MCMC.params$n.burnin,
             n.thin = MCMC.params$n.thin,
             n.iter = MCMC.params$n.samples,
             DIC = T, 
             parallel=T)
  
  saveRDS(jm, file = "RData/JAGS_out_Dist_rSlope_rInt_3Covs_Aug2020.rds")
  
} else {
  jm <- readRDS("RData/JAGS_out_Dist_rSlope_rInt_3Covs_Aug2020.rds")
}

summary(jm)
```

At least they converged fine! 


```{r}
summary.df <- data.frame(jm$summary) %>%
  rownames_to_column("Parameter")

# find the beach-specific growth rate in percent
#mcmc_trace(jm.MS$samples, pars = "a1pc[38]")

summary.df %>% filter(str_detect(Parameter, pattern = "a1pc")) %>%
  rownames_to_column("beach") %>%
  transmute(ID = as.numeric(beach),
            mean_a1pc = mean,
            sd_a1pc = sd,
            low25_a1pc = X2.5.,
            median_a1pc = X50.,
            upper975_a1pc = X97.5.)-> a1pc

a1pc %>% left_join(lat.dat, by = "ID") %>%
  arrange(latc) -> a1pc

summary.df %>% 
  filter(str_detect(Parameter, pattern = "mu.a1")) %>%
  rownames_to_column("beach") %>%
  transmute(ID = as.numeric(beach),
            mean_mu.a1 = mean,
            sd_mu.a1 = sd,
            low25_mu.a1 = X2.5.,
            median_mu.a1 = X50.,
            upper975_mu.a1 = X97.5.)-> mu.a1

mu.a1 %>% left_join(lat.dat, by = "ID") %>%
  arrange(latc) -> mu.a1

#head(a1pc.2008)
```


Look at why a1pc.2008 is quite different from a1pc.new...

Compare results visually...
```{r}

ggplot() + 
  geom_point(data = a1pc, 
             aes(x = beach, y = mean_a1pc),
             color = "blue") + 
  geom_errorbar(data = a1pc,
                aes(ymin = low25_a1pc,
                    ymax = upper975_a1pc,
                    x = beach),
                color = "blue") 
```

TAke a look at the overdispersion variance parameter.

```{r}
mcmc_dens(jm$samples, pars = "sigma.e")
```

It's definitely not zero... 



