---
title: "R Notebook"
output: html_notebook
---

Trend analysis of leatherback turtles nesting in Florida, Puerto Rico, and St Croix with Kelly Stewart.

New data were received in August 2020. 

```{r}
rm(list=ls())
library(tidyverse)
library(ggplot2)
library(readr)
library(lubridate)
library(jagsUI)
library(bayesplot)
library(stringr)

source("Caribbean_Dc_fcns.R")

MCMC.params <- list(n.samples = 75000,
                    n.burnin = 45000,
                    n.thin = 5,
                    n.chains = 5)

MCMC.params$n.per.chain = (MCMC.params$n.samples - MCMC.params$n.burnin)/MCMC.params$n.thin

```

Get data;

```{r}
col.def <- cols(ID = col_integer(),
                year = col_integer(),
                beach = col_character(),
                latitude = col_double(),
                distance = col_double(),
                days_week = col_integer(),
                days_year = col_integer(),
                nests = col_integer())

STX.nest.counts <- read_csv("data/STX_Sept2020.csv", 
                           col_types = col.def) %>% 
  mutate(beach_f = as.factor(toupper(beach)),
         dataset = "STX",
         ID2 = as.numeric(as.factor(ID)))


```


STX

```{r}
STX.nest.counts %>% 
  group_by(ID2) %>%
  summarise(n = n()) -> STX.ns

STX.nest.counts %>% 
  select(ID2, ID, beach_f, year) %>% 
  group_by(ID) %>%
  summarise(ID2 = first(ID2),
            name = first(beach_f),
            median.yr = median(year),
            min.yr = min(year),
            n = n()) %>%
  mutate(beach = ID,
         ID2 = ID2,
         name = name) -> STX.lat.dat

median.yr <- select(STX.lat.dat, c(ID, median.yr, min.yr)) %>%
  right_join(STX.nest.counts, by = "ID") %>% 
  select(ID, median.yr, min.yr, year)

# changed from centering to left-adjusted for years. I think t >= 0 makes more
# sense explaining the intercept than centering time. With t >= 0, it is the 
# log(mu) at time 0 for each beach, whereas if time is centered, it is log(mu) at
# some arbitrary year for each beach. 

STX.jags.data <- list(N = length(STX.nest.counts$nests),
                      nbeach = length(unique(STX.nest.counts$ID2)),
                      count = STX.nest.counts$nests,
                      beach = STX.nest.counts$ID2,
                      yearc = median.yr$year - median.yr$min.yr)

data.vector <- STX.jags.data$count %>% 
  rep(each = MCMC.params$n.chains * MCMC.params$n.per.chain)

parameters <- c("a0", "a1", "beta", 
                "sigma.e", "deviance",
                "Devobs", "Devpred", "loglik")

```


```{r}
# sampling happens every day so Days of week are constant 
model.names <- c("1Cov", "0Cov")
out.names <- c("DayYr", "0Cov") 
X.STX <- list(STX.nest.counts$days_year - median(STX.nest.counts$days_year),
              0)

#jm <- list(length = length(out.names))
loo.out <- list()
Rmax <- vector(mode = "numeric", length = length(out.names))

for (k in 1:length(out.names)){
  MCMC.params$model.file = paste0("models/Model_JAGS_Pois_", model.names[k], "_NoLat.txt")
  STX.jags.data$X <- X.STX[[k]]

  if (!file.exists(paste0("RData/JAGS_out_Pois_", out.names[k], "_NoLat_STX.rds"))){
    jm <- jags(data = STX.jags.data,
               #inits = inits,
               parameters.to.save= parameters,
               model.file = MCMC.params$model.file,
               n.chains = MCMC.params$n.chains,
               n.burnin = MCMC.params$n.burnin,
               n.thin = MCMC.params$n.thin,
               n.iter = MCMC.params$n.samples,
               DIC = T, 
               parallel=T)
    
    saveRDS(jm, 
            file = paste0("RData/JAGS_out_Pois_", out.names[k], "_NoLat_STX.rds"))
    
  } else {
    jm <- readRDS(file = paste0("RData/JAGS_out_Pois_", out.names[k], "_NoLat_STX.rds"))
  }

  Rmax[k] <- max(unlist(lapply(jm$Rhat, FUN = max)))
  
  if (!file.exists(paste0("RData/LOOIC_Pois_", out.names[k], "_NoLat_STX.rds"))){
    loo.out[[k]] <- compute.LOOIC(loglik = jm$sims.list$loglik, 
                                  data.vector = data.vector, 
                                  MCMC.params = MCMC.params)
    saveRDS(loo.out[[k]], 
            file = paste0("RData/LOOIC_Pois_", out.names[k], "_NoLat_STX.rds"))
    
  } else {
    loo.out[[k]] <- readRDS(file = paste0("RData/LOOIC_Pois_", out.names[k], "_NoLat_STX.rds"))
  }
  
}

Rmax
```


At least they converged fine! But not a great fit according to Pareto k diagnostic values. I tried the second order polynomial but didn't improve the fit much... Maybe this is where negative binomial is needed? 

```{r}
loo.out
```


```{r}
parameters <- c("a0", "a1", "beta", "r",
                "sigma.e", "deviance",
                "Devobs", "Devpred", "loglik")

#jm <- list(length = length(out.names))
loo.out <- list()
Rmax <- vector(mode = "numeric", length = length(out.names))

for (k in 1:length(out.names)){
  MCMC.params$model.file = paste0("models/Model_JAGS_negbin_", model.names[k], "_NoLat.txt")
  STX.jags.data$X <- X.STX[[k]]

  if (!file.exists(paste0("RData/JAGS_out_negbin_", out.names[k], "_NoLat_STX.rds"))){
    jm <- jags(data = STX.jags.data,
               #inits = inits,
               parameters.to.save= parameters,
               model.file = MCMC.params$model.file,
               n.chains = MCMC.params$n.chains,
               n.burnin = MCMC.params$n.burnin,
               n.thin = MCMC.params$n.thin,
               n.iter = MCMC.params$n.samples,
               DIC = T, 
               parallel=T)
    
    saveRDS(jm, 
            file = paste0("RData/JAGS_out_negbin_", out.names[k], "_NoLat_STX.rds"))
    
  } else {
    jm <- readRDS(file = paste0("RData/JAGS_out_negbin_", out.names[k], "_NoLat_STX.rds"))
  }

  Rmax[k] <- max(unlist(lapply(jm$Rhat, FUN = max)))
  
  if (!file.exists(paste0("RData/LOOIC_negbin_", out.names[k], "_NoLat_STX.rds"))){
    loo.out[[k]] <- compute.LOOIC(loglik = jm$sims.list$loglik, 
                                  data.vector = data.vector, 
                                  MCMC.params = MCMC.params)
    saveRDS(loo.out[[k]], 
            file = paste0("RData/LOOIC_negbin_", out.names[k], "_NoLat_STX.rds"))
    
  } else {
    loo.out[[k]] <- readRDS(file = paste0("RData/LOOIC_negbin_", out.names[k], "_NoLat_STX.rds"))
  }
  
}

Rmax
```

Using negative binomial seems to result in better Pareto k values. LOOIC indicated the model with no covariate was a better one, where logit(p[i]) = a0 + a1 * year + e.

```{r}
loo.out
```


But, it doesn't make sense to have just one slope for the time term... Look at the pareto k values

```{r}
STX.paretok.data <- data.frame(count = STX.jags.data$count,
                            yearc = STX.jags.data$yearc,
                            year = STX.nest.counts$year,
                            pareto_k = loo.out[[2]]$loo.out$diagnostics$pareto_k)

ggplot(data = STX.paretok.data) +
  geom_point(aes(x = yearc, 
                 y = count,
                 color = pareto_k)) +
  scale_color_continuous(type = "viridis")


```

Pareto k values are high at around the inflection point in abundance trend... so, we need to consider a more flexible function that takes into account there is a large change in slope from positive to negative. 

Poisson

```{r}

parameters <- c("a0.1", "a1.1", "a0.2", "a1.2",
                "beta", "r", "year.change",
                "sigma.e", "deviance",
                "Devobs", "Devpred", "loglik")

STX.jags.data$maxT <- 25
STX.jags.data$minT <- 15

#jm <- list(length = length(out.names))
loo.out <- list()
Rmax <- vector(mode = "numeric", length = length(out.names))

for (k in 1:length(out.names)){
  MCMC.params$model.file = paste0("models/Model_JAGS_Pois_2slopes_", model.names[k], "_NoLat.txt")
  STX.jags.data$X <- X.STX[[k]]
  
  if (!file.exists(paste0("RData/JAGS_out_Pois_2slopes_", out.names[k], "_NoLat_STX.rds"))){
    jm <- jags(data = STX.jags.data,
               #inits = inits,
               parameters.to.save= parameters,
               model.file = MCMC.params$model.file,
               n.chains = MCMC.params$n.chains,
               n.burnin = MCMC.params$n.burnin,
               n.thin = MCMC.params$n.thin,
               n.iter = MCMC.params$n.samples,
               DIC = T, 
               parallel=T)
    
    saveRDS(jm, 
            file = paste0("RData/JAGS_out_Pois_2slopes_", out.names[k], "_NoLat_STX.rds"))
    
  } else {
    jm <- readRDS(file = paste0("RData/JAGS_out_Pois_2slopes_", out.names[k], "_NoLat_STX.rds"))
  }

  Rmax[k] <- max(unlist(lapply(jm$Rhat, FUN = max)))
  
  if (!file.exists(paste0("RData/LOOIC_Pois_2slopes_", out.names[k], "_NoLat_STX.rds"))){
    loo.out[[k]] <- compute.LOOIC(loglik = jm$sims.list$loglik, 
                                  data.vector = data.vector, 
                                  MCMC.params = MCMC.params)
    saveRDS(loo.out[[k]], 
            file = paste0("RData/LOOIC_Pois_2slopes_", out.names[k], "_NoLat_STX.rds"))
    
  } else {
    loo.out[[k]] <- readRDS(file = paste0("RData/LOOIC_Pois_2slopes_", out.names[k], "_NoLat_STX.rds"))
  }
  
}

Rmax
```

I can't make these converge. Poisson may be not useful? 

Negative binomial

```{r}
parameters <- c("a0.1", "a1.1", "a0.2", "a1.2",
                "beta", "r", "year.change", "mu",
                "sigma.e", "deviance",
                "Devobs", "Devpred", "loglik")

#jm <- list(length = length(out.names))
loo.out <- list()
Rmax <- vector(mode = "numeric", length = length(out.names))
STX.jags.data$maxT <- 25
STX.jags.data$minT <- 15

for (k in 1:length(out.names)){
  MCMC.params$model.file = paste0("models/Model_JAGS_negbin_2slopes_", model.names[k], "_NoLat.txt")
  STX.jags.data$X <- X.STX[[k]]

  if (!file.exists(paste0("RData/JAGS_out_negbin_2slopes_", out.names[k], "_NoLat_STX.rds"))){
    jm <- jags(data = STX.jags.data,
               #inits = inits,
               parameters.to.save= parameters,
               model.file = MCMC.params$model.file,
               n.chains = MCMC.params$n.chains,
               n.burnin = MCMC.params$n.burnin,
               n.thin = MCMC.params$n.thin,
               n.iter = MCMC.params$n.samples,
               DIC = T, 
               parallel=T)
    
    saveRDS(jm, 
            file = paste0("RData/JAGS_out_negbin_2slopes_", out.names[k], "_NoLat_STX.rds"))
    
  } else {
    jm <- readRDS(file = paste0("RData/JAGS_out_negbin_2slopes_", out.names[k], "_NoLat_STX.rds"))
  }

  Rmax[k] <- max(unlist(lapply(jm$Rhat, FUN = max)))
  
  if (!file.exists(paste0("RData/LOOIC_negbin_2slopes_", out.names[k], "_NoLat_STX.rds"))){
    loo.out[[k]] <- compute.LOOIC(loglik = jm$sims.list$loglik, 
                                  data.vector = data.vector, 
                                  MCMC.params = MCMC.params)
    saveRDS(loo.out[[k]], 
            file = paste0("RData/LOOIC_negbin_2slopes_", out.names[k], "_NoLat_STX.rds"))
    
  } else {
    loo.out[[k]] <- readRDS(file = paste0("RData/LOOIC_negbin_2slopes_", out.names[k], "_NoLat_STX.rds"))
  }
  
}

Rmax
```

At least these converge okay. Well... when I make the prior for the r parameter wider (gamma(2, 0.5) to gamma(2, 0.1)), convergence becomes an issue. So, the model is inappropriate. Increasing the number of samples helped converge with gamma(2, 0.1) but posteriors don't look great

```{r}
mcmc_trace(jm$samples, c("a0.1", "a1.1", "a0.2", "a1.2"))
```

It doesn't make any sense that the intercepts are about the same for those two conditions... What am I doing wrong here? 




